---
title: "NFL Wins Pool Update"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: ""
date: ""
params:
  week: 2
output: 
  html_document:
    preserve_yaml: FALSE
---

```{r setup, include=FALSE, warning = FALSE}

# Instructions for GIT Upload:
# 1) hope RStudio is still logged into GIT (don't remember password or process)
# 2) knit document to generate HTML output
# 3) On GIT icon on top bar, select 'Commit'
# 4) in top-left pane, select 'index.html' (knitted html file) - usually at the bottom
# 5) in top-right pane, type commit message (usually '[year] Week [x] update')
# 6) click 'commit'; new window will pop up; close when process is complete
# 7) click 'push' in top-right corner; new window will pop up; close when process is complete
# all done

knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(forcats)
library(rvest)
library(formattable)
library(knitr)
library(kableExtra)
library(fuzzyjoin)
library(readxl)
library(ggplot2)
library(plotly)
library(nflfastR)
```

```{r read files, echo= FALSE, warning = F}

Draft <- read.csv("DraftResults25.csv", stringsAsFactors = F)
Draft$Team <- factor(Draft$Team, levels = Draft$Team[1:10])
#
# Read NFL Standings
url <- "http://www.espn.com/nfl/standings/_/season/2025/group/league"
webpage <- read_html(url)

teams <- webpage %>% html_nodes(".hide-mobile a") %>% html_text
wins <- webpage %>% html_nodes(".Table__TD:nth-child(1) .stat-cell") %>% html_text %>% as.numeric()
losses <- webpage %>% html_nodes(".Table__TD:nth-child(2) .stat-cell") %>% html_text %>% as.numeric()
ties <- webpage %>% html_nodes(".Table__TD:nth-child(3) .stat-cell") %>% html_text %>% as.numeric()


nfl.standings <- data.frame(Team = teams, W = wins, L = losses, T = ties) %>%
  mutate(GP = rowSums(.[2:4]))

# Determine who each team plays in current week

#Schedule_Key <- data.frame(Week = seq(1:22), Date = seq(from = as.Date("2022-09-08"), to = as.Date("2023-02-02"), by = #7))
#Current.Week <- Schedule_Key %>% filter(Date >= Sys.Date()) %>% slice(1) %>% pull(Week) 
Schedule <- week_sched <- nflreadr::load_schedules(seasons = 2025) %>% filter(week == params$week)

Schedule_Week <- rbind(
  Schedule %>% select(TEAM = home_team, Opponent = away_team, spread = spread_line),
  Schedule %>% mutate(spread_line = spread_line * -1) %>% select(TEAM = away_team, Opponent = home_team, spread = spread_line)
) %>%
  arrange(TEAM) %>%
  left_join(
    nflreadr::load_teams() %>%
      select(team_abbr, Merge = team_name),
    by = c("TEAM" = "team_abbr")
  )

#Schedule_Week$Merge <- sort(as.character(nfl.standings$Team))[c(1:16,19, #18,17,20:27,29,28,30:32)]

Num.Byes <- sum(Schedule_Week$Opponent == "BYE")/2


# get order for merge
order_db <- data.frame(Merge = sort(as.character(nfl.standings$Team)),
                       stringsAsFactors = F) %>%
  mutate(S = seq(1:n()),
         Name = trimws(gsub("^.+ *( .+$)", "\\1", Merge))) %>%
  arrange(Name)

# Read ESPN FPI projections
url.4 <- "https://www.espn.com/nfl/fpi/_/view/projections"
webpage <- read_html(url.4)

teams.espn <- webpage %>% html_nodes("td:nth-child(1)") %>% html_text
teams.espn <- teams.espn[1:32]

record <- webpage %>% html_nodes("td:nth-child(2)") %>% html_text
espn.wins <- as.numeric(gsub("-.+$", "", record))

Projections.espn <- data.frame(Merge = as.character(teams.espn), Proj.W = espn.wins) %>% arrange(Merge)
Projections.espn$Merge <- as.character(Projections.espn$Merge)

Merge.Key <- data.frame(Short = sort(Schedule_Week$TEAM)[c(1:16,18:19,17,20:32)],
                          Abbrev = (order_db %>% arrange(Merge) %>% pull(Name))[c(1:16,19,17:18,20:27,29,28,30:32)],
                          Full = (order_db %>% arrange(Merge) %>% pull(Merge))[c(1:16,19,17:18,20:27,29,28,30:32)]) %>%
  mutate(First = trimws(gsub("(^.+ *) .+$", "\\1", Full)))


# Read vegas odds
t <- format(Sys.Date(), "%B %d, %Y")

```

Last updated `r t`

## Standings

```{r echo = F, warning=F}
pool.standings <- left_join(Draft[,3:4], nfl.standings, by = c("Selection" = "Team")) %>%
  left_join(Projections.espn[,1:2], by = c("Selection" = "Merge")) %>%
  arrange(Selection) %>%
  group_by(Team) %>%
  summarise_at(vars(-Selection), funs(sum)) %>%
  mutate(Rank = rank(-W, ties.method = "min"),
         Games.Left = 51 - GP,
         W.Pct = percent(W/GP, digits = 1),
         W.Pct.Rk = rank(-W.Pct, ties.method = "min"),
         Proj.Rank_FPI = rank(-Proj.W, ties.method = "min"),
         ) %>%
  arrange(Rank) %>%
  select(Rank, Team, W, Games.Left, W.Pct, W.Pct.Rk, Proj.W, Proj.Rank_FPI, #Proj.Rank_538
         ) %>%
  rename(Wins = W) 

pool.standings %>%
  kable(col.names = c("Rank", "Team", "Wins", "Remaining Games", "Win %", "Win % Rank", "Projected Wins (FPI)", "Projected Rank (FPI)"), 
        align = "c") %>%
#  kable(col.names = c("Rank", "Team", "Wins", "Remaining Games", "Win %"), align = "c") %>%
  kable_styling(bootstrap_options = c("hover","condensed", "responsive"), full_width = F)

```

* Projected wins taken from [ESPN FPI projections](https://www.espn.com/nfl/fpi/_/view/projections).

## Expanded Standings

```{r echo = F, warning = F}

expanded.standings <- left_join(nfl.standings[, 1:2], Draft[, 3:4], by = c("Team" = "Selection")) %>%
  rename(Picker = Team.y) %>%
  mutate(Picker = as.character(Picker)) %>%
  filter(!is.na(Picker)) %>%
  left_join(Schedule_Week[, 2:4], by = c("Team" = "Merge")) %>%
  left_join(pool.standings[, 2:3], by = c("Picker" = "Team")) %>%
  left_join(Projections.espn, by = c("Team" = "Merge")) %>%
  mutate(Prob = 1-round(1/(10^((25*spread)/400)+1),2)) %>%
  arrange(desc(Wins), Picker, desc(W), Team)

Pool.Order <- unique(expanded.standings$Picker)

names_broken <- c("Team", "Wins","Projected<br/>Wins" ,"Next<br/>Opponent", "Win Probability<br/>in Next Game")

expanded.standings %>%
  select(Team, W, Proj.W, Opponent, Prob) %>%
  mutate(Prob = ifelse(is.na(Prob), "", Prob)) %>%
  kable(col.names = names_broken, escape = F) %>%
  kable_styling(bootstrap_options = c("hover"), full_width = F, position = "left") %>%
  group_rows(Pool.Order[1], 1,3) %>%
  group_rows(Pool.Order[2], 4,6) %>%
  group_rows(Pool.Order[3], 7,9) %>%
  group_rows(Pool.Order[4], 10,12) %>%
  group_rows(Pool.Order[5], 13,15) %>%
  group_rows(Pool.Order[6], 16,18) %>%
  group_rows(Pool.Order[7], 19,21) %>%
  group_rows(Pool.Order[8], 22,24) %>%
  group_rows(Pool.Order[9], 25,27) %>%
  group_rows(Pool.Order[10], 28,30) 
```

* Win probability calculated from [covers.com odds](https://www.covers.com/nfl/week-2-odds)

## Weekly Head-to-Heads

```{r H2H, echo = F, warning=F}
First_Place <- max(pool.standings$Wins)
Third_Place <- pool.standings$Wins[3]

# fix road team for london game
# > none for 2020; keeping code for the future
#Schedule_Long$Opponent[Schedule_Long$TEAM == "CAR" & Schedule_Long$Week == 6] <- "@TB"
#Schedule_Long$Opponent[Schedule_Long$Opponent == "LAR" & Schedule_Long$Week == 8] <- "@LAR"
#Schedule_Long$Opponent[Schedule_Long$Opponent == "LAC" & Schedule_Long$Week == 11] <- "@LAC"

#Matchups <- Schedule_Long %>%
#  filter(Week == Current.Week,
#         grepl("@", Opponent)) %>%
#  select(-Week) %>%
#  mutate(Opponent = gsub("@", "", Opponent),
#         Game = seq(1:n())) %>%
#  rename(Away = TEAM,
#         Home = Opponent) %>%
#  gather("Location", "Team", Away:Home) %>%
#  mutate(Team = ifelse(Team == "RAM", "LAR", Team)) %>%
#  left_join(Merge.Key[,c(1,3)], by = c("Team" = "Short")) %>%
#  left_join(Draft[,c(3,4)] %>% rename("Person" = "Team"), by = c("Full" = "Selection")) %>%
#  left_join(pool.standings[, c(2:3)] %>% rename("Person" = "Team"), by ="Person") %>%
#  left_join(Spreads[, c(3,4)], by = "Full") %>%
#  select(-Full) %>%
#  unite("temp", Team,Person,Wins,Prob, sep = "-") %>%
#  spread(Location, temp) %>%
#  separate(Home, c("Home", "Home_Person", "Home_Wins", "Home_Prob"), sep = "-") %>%
#  separate(Away, c("Away", "Away_Person", "Away_Wins", "Away_Prob"), sep = "-") %>%
#  mutate_at(vars(Home_Wins, Home_Prob, Away_Wins, Away_Prob), funs(as.numeric(.))) %>%
#  mutate_at(vars(Home_Wins, Away_Wins), funs(ifelse(is.na(.), min(pool.standings$Wins), .))) %>%
#  rowwise() %>%
#  mutate(Rank.Rating_First = (First_Place - Home_Wins) + (First_Place - Away_Wins),
#         Rank.Rating_Money = (max(0, Third_Place - Home_Wins, na.rm = T) + max(0, Third_Place - #Away_Wins, na.rm = T)),
#         Prob.Rating = max(Home_Prob, Away_Prob)) %>%
#  ungroup() %>%
#  arrange(Rank.Rating_First, Rank.Rating_Money, Prob.Rating) %>%
#  mutate(Leverage.Score = 1:n(),
#         Output = paste0("(", Away_Person,") ", Away, " @ ", Home, " (", Home_Person,")"),
#         At = "@") %>%
#  select(Home_Person, Home, At, Away, Away_Person, Leverage.Score) %>%
#  mutate_at(vars(Away_Person, Home_Person), funs(ifelse(. == "NA", "", .))) %>%
#  filter(Home_Person != Away_Person)
#
#colfunc <- colorRampPalette(c("red", "white"))

#Matchups %>%
#  mutate(
#    Indicator.Cell = "ReallyLongName",
#    Indicator.Cell = cell_spec(
#    Indicator.Cell, color = colfunc(nrow(Matchups)), 
#    background = colfunc(nrow(Matchups))
#  )) %>%
#  select(-Leverage.Score) %>%
#    kable(col.names = c("Person", "Team", "", "Team", "Person", "Leverage"), escape = F) %>%
#  kable_styling(bootstrap_options = c("hover"), full_width = F, position = "left")

matchup_db <- Schedule %>%
  mutate(Game = seq(1:n())) %>%
  select(Game, Away = away_team, Home = home_team, spread = spread_line) %>%
  gather("Location", "Team", Away:Home) %>%
  left_join(nflreadr::load_teams() %>%
              select(team_abbr, team_name),
            by = c("Team" = "team_abbr")) %>%
  left_join(Draft[,c(3,4)] %>% rename("Person" = "Team"), by = c("team_name" = "Selection")) %>%
  left_join(pool.standings[, c(2:3)] %>% rename("Person" = "Team"), by ="Person") %>%
  mutate(Prob = 1-round(1/(10^((25*spread)/400)+1),2)) %>%
  select(-team_name) %>%
  unite("temp", Team,Person,Wins,Prob, sep = "-") %>%
  spread(Location, temp) %>%
  separate(Home, c("Home", "Home_Person", "Home_Wins", "Home_Prob"), sep = "-") %>%
  separate(Away, c("Away", "Away_Person", "Away_Wins", "Away_Prob"), sep = "-") %>%
  mutate(across(.cols = c(Home_Wins, Home_Prob, Away_Wins, Away_Prob), .fns = as.numeric)) %>%
  mutate(across(.cols = c(Home_Wins, Away_Wins), .fns = ~ ifelse(is.na(.x), min(pool.standings$Wins), .x))) %>%
  mutate(across(.cols = c(Home_Wins, Away_Wins), .fns = ~ First_Place - .)) %>%
  group_by(Game) %>%
  mutate(min_prob = min(Home_Prob, Away_Prob),
         label = glue::glue("{Away} ({Away_Person}) @ {Home} ({Home_Person}) - {min_prob*100}%"))

p <- ggplot(matchup_db, aes(x = Home_Wins, y = Away_Wins, size = min_prob, text = label)) +
  geom_point(position = position_jitter(width = 0.1, height = 0.1), shape = 21, color = "red", fill = "blue", alpha = 0.25) +
  scale_x_reverse(name = "Player 1 distance from 1st place") +
  scale_y_reverse(name = "Player 2 distance from 1st place")

ggplotly(p, tooltip = "label")

```
Leverage measured by:  
* Circles to the top right are games between people closest to first place  
* Bigger circles indicate games that are more likely to be closely contested

## Strength of Schedule in the coming weeks

```{r sos, echo = F, warning = F}

team_list <- Draft %>%
  left_join(nflreadr::load_teams() %>%
              select(team_abbr, team_name),
            by = c("Selection" = "team_name")) %>%
  select(Team, team_abbr)
  

sched_upcoming <- nflreadr::load_schedules(seasons = 2024) %>% 
  filter(between(week, params$week, min(params$week + 3, max(week)))) %>%
  nflreadr::clean_homeaway() %>%
  select(game_id, week, team, opponent) %>%
  left_join(nflreadr::load_teams() %>%
              select(team_abbr, team_name),
            by = c("opponent" = "team_abbr")) %>%
  left_join(Projections.espn, by = c("team_name" = "Merge")) %>%
  select(game_id, week, team, opponent, Proj.W) %>%
  left_join(nflreadr::load_teams() %>%
              select(team_abbr, team_name),
            by = c("team" = "team_abbr")) %>%
  left_join(Draft[,c(1,3,4)] %>% rename("Person" = "Team"), by = c("team_name" = "Selection")) %>%
  filter(!is.na(Person)) %>%
  select(Person, team, Round, week, opponent, Proj.W) %>%
  rowwise() %>%
  mutate(self_game = opponent %in% team_list$team_abbr[team_list$Team == Person]) %>%
  group_by(Person) %>%
  mutate(self_games = sum(self_game) / 2) %>%
  filter(!self_game) %>%
  mutate(avg_sos = median(Proj.W),
         tooltip = paste0("Week ", week, ": ", team, " against ", opponent)) 
  
p <- sched_upcoming %>%
  ggplot(aes(x = fct_reorder(Person, -avg_sos), y = Proj.W, text = tooltip)) +
  geom_jitter(aes(color = factor(Round)), size = 2, width = 0.1) +
  geom_boxplot(fill = NA, color = "grey80") +
  geom_text(aes(y = min(sched_upcoming$Proj.W) - 1, label = self_games), size = 3) +
  coord_flip() +
  labs(x  = "", y = "Opponent projected wins") +
  ggtitle(paste0("Strength of schedule over next ", length(unique(sched_upcoming$week)), " weeks")) +
  scale_colour_manual(name = "Team #", values = c("1", "2", "3")) +
  annotate("text", x= 0.5 , y= min(sched_upcoming$Proj.W) - 1, label= "Self-games", size = 3, vjust = 0) +
  theme_minimal()

w <- ggplotly(p, tooltip = "tooltip")

w %>%
  style(hoverinfo = "skip", traces = 4)


```


## Draft Analysis

```{r echo = F, warning = F}
full.standings <- left_join(nfl.standings[, 1:2], Draft[, 2:4], by = c("Team" = "Selection")) %>%
  rename(Picker = Team.y) %>%
  mutate(Picker = as.character(Picker))

full.standings$Pick[is.na(full.standings$Pick)] <- c(31,32)
full.standings$Picker[is.na(full.standings$Picker)] <- "Not Picked"

pool.rosters <- left_join(Draft, nfl.standings, by = c("Selection" = "Team")) %>%
  mutate(Exp.W = nfl.standings$W[Pick],
         PvE = W - Exp.W) %>%
  group_by(Team) %>%
  mutate(Total = sum(W)) %>%
  rowwise() %>%
  mutate(better.pick = sum(full.standings$Pick[full.standings$Picker != Team] > Pick & full.standings$W[full.standings$Picker != Team] > W)) %>%
  arrange(desc(Total), Team, Pick)


names_broken <- c("Team", "Wins", "Draft<br/>Spot", "Expected<br/>Wins", "Performance vs.<br/>Expectation", "Better<br/>Choices")

pool.rosters %>%
  ungroup() %>%
  select(Selection, W, Pick, Exp.W, PvE, better.pick) %>%
  kable(col.names = names_broken, escape = F) %>%
  kable_styling(bootstrap_options = c("hover"), full_width = F, position = "left") %>%
  group_rows(pool.standings$Team[1], 1,3) %>%
  group_rows(pool.standings$Team[2], 4,6) %>%
  group_rows(pool.standings$Team[3], 7,9) %>%
  group_rows(pool.standings$Team[4], 10,12) %>%
  group_rows(pool.standings$Team[5], 13,15) %>%
  group_rows(pool.standings$Team[6], 16,18) %>%
  group_rows(pool.standings$Team[7], 19,21) %>%
  group_rows(pool.standings$Team[8], 22,24) %>%
  group_rows(pool.standings$Team[9], 25,27) %>%
  group_rows(pool.standings$Team[10], 28,30)



```



## Draft Results

```{r draft, echo=FALSE}
graph_data <- full.standings %>%
  select(Picker, Team, W, Pick)

graph_data$Picker <- factor(graph_data$Picker, levels = c(as.character(Draft$Team[1:10]), "Not Picked"))

ggplot(graph_data, aes(x = Picker, y = Pick)) + 
  geom_point() +
  geom_text(aes(label = Team, color = W), size = 3, hjust = 0, nudge_x = 0.15) + 
  scale_y_reverse(name = "Overall Selection", breaks = c(1,10,20,30)) +
  scale_color_gradient(name = "Wins", low = "red", high = "darkblue")+
  coord_cartesian(xlim = c(0.75,14)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


